Nonlinear transformations and bias correction

### 2021-09-03 logist_normal_bias.rmd

* Although it might not matter, we do need pick a reference point to start from (e.g. the prediction for the model center). Then, *on the linear predictor scale*, we can compute the derivative (which is just the coefficient for the focal parameter unless there are interactions, in which case we need to incorporate the conditional values/distribution of the non-focal parameters in the calculation). If there are no interactions, then the derivative on the response scale is the same for the entire distribution and by the chain rule is the product of the derivative ($\beta_f = d\eta/dx_f$) and the derivative of the inverse link function at the point ($dg^{-1}(\eta)/d\eta$, which is the `mu.eta` component of a `family` object in R). How does variability (due to variation in the non-focal parameters + interactions) affect this result?? Suppose $\hat \beta_f(x_n)$ is the derivative for a particular set of non-focal parameters (i.e. $\hat \beta f(x_n) \equiv \beta_f + \sum \beta_{f\nset{i}} \nset{x}$ where the second term is the sum of the relevant interaction coefficients). Then I claim that the marginal effect at a particular value of the focal parameter $\xfprime$ is

@@

- compute $\yref$ by taking $\cdist$ at the center value of the focal parameter and finding the mean (either directly or by computing the mean and variance of $\eta$ and using the logistic-normal calculation
- for each step away from the center
   - find $\cdist$ for the new value of $x_f$;
   - if necessary (i.e. if the model has interactions with the focal parameter) compute $\hat \beta f$ across $\cdist$;
   - compute the derivative across $\cdist$
   - compute the mean value of $\phi = dg^{-1}/d\eta \hat \beta f$
   - add $\phi \Delta x_f$ to the effects-plot value

>>>>

* Although it might not matter, we do need pick a reference point to start from (e.g. the prediction for the model center). Then, *on the linear predictor scale*, we can compute the derivative (which is just the coefficient for the focal parameter unless there are interactions, in which case we need to incorporate the conditional values/distribution of the non-focal parameters in the calculation). If there are no interactions, then the derivative on the response scale is the same for the entire distribution and by the chain rule is the product of the derivative ($\beta_f = d\eta/dx_f$) and the derivative of the inverse link function at the point ($dg^{-1}(\eta)/d\eta$, which is the `mu.eta` component of a `family` object in R). How does variability (due to variation in the non-focal parameters + interactions) affect this result?? Suppose $\hat \beta_f(x_n)$ is the derivative for a particular set of non-focal parameters (i.e. $\hat \beta_f(x_n) \equiv \beta_f + \sum \beta_{f\nset{i}} \nset{x}$ where the second term is the sum of the relevant interaction coefficients). Then I claim that the marginal effect at a particular value of the focal parameter $\xfprime$ is

@@

- compute $\yref$ by taking $\cdist$ at the center value of the focal parameter and finding the mean (either directly or by computing the mean and variance of $\eta$ and using the logistic-normal calculation
- for each step away from the center
   - find $\cdist$ for the new value of $x_f$;
   - if necessary (i.e. if the model has interactions with the focal parameter) compute $\hat \beta_f$ across $\cdist$;
   - compute the derivative across $\cdist$
   - compute the mean value of $\phi = dg^{-1}/d\eta \hat \beta_f$
   - add $\phi \Delta x_f$ to the effects-plot value
